---
title: "Initial Korean covid 19"
author: "Galilea  Hernandez"
date: "09/26/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  
```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

After observing the data in the graph, the data shows that although there were about 1800 females to 1400 males that contracted covid, twice as many males died form covid in 2020. The percentage of females that died was about 35% while the percentage of males was well over 60%.
  





  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

From the numerical results, 56% of the females remained alive, while 445 of the males died. This leads to 36% of females dying while 64% of males died. When observing the totals from a 100.0, 98.7/100 females lived and 1.4/100 died. 97/100 males lived, and 3/100 males died.

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Three tests were performed to identify if dying from covid is independant of a patients sex.The p-value for the first test 0.001, for the second test the p-value is 0.0012 and for the third test the p-value is 9e-04 indicating the difference in statistical probability of female patients versus male patients in terms of how many died from covid.

## Conclusions

To observe whether patients in korea who contracted covid died independant of sex we graphed collected data in a barplot, then scripted a numerical summary to show the percentages of the male vs feemales patients that passed. Lastly, we obtained inferential data by running three tests including Pearsons CHi test, Chi square test, and a fisher test.

After carefully analyzing the data from all out data outputs, we saw that the females who contracted covid vs ales that contrated covid. males were twice as probable of dying than females.